---
title: "House of Lords Petitions Explorer"
author: "Sharon Howard"
categories: [HoL]
execute: 
  warning: false
  message: false
  echo: false 
---


An interactive table for exploring the Power of Petitioning [House of Lords petitions](https://www.british-history.ac.uk/petitions/house-of-lords) data. 


```{r}
source(here::here("_R/data.R"))

source(here::here("_R/reactable.R"))

#library(reactablefmtr) # mainly additional themes


# sticky column styles
# put these in reactable.R. with various notes. 
# added to reactable.R: headerMakeSpace 
# js function to replace _ in header names with space. maybe should do camelCase as well but that's harder and i don't often use it. 
```


```{r}
bho_hl_urls <-
tribble(~bho_path,
  "1597", 	
"1601", 	
"1606", 	
"1610", 	
"1613", 	
"1614", 	
"1619", 	
"1620", 	
"1621", 	
"1624", 	
"1640", 	
"1648", 	
"1661", 	
"1671", 	
"1679", 	
"1689", 	
"1696") |>
  mutate(year = parse_number(bho_path)) |>
  mutate(bho_path= paste0("house-of-lords/", bho_path))


data_reactable <-
hol_sqlite |>  
  select(petition_id, petitioner, year, date, themes, petition_type, petition_gender, description, reference) |>
  mutate(reference = paste("PA", reference)) |>
  mutate(themes = str_replace_all(themes, " *\\| *", ", ")) |>
  mutate(petition_gender = case_when(
    petition_gender=="f" ~ "female", 
    petition_gender=="fm" ~ "mixed", 
    petition_gender=="m" ~ "male", 
    petition_gender=="u" ~ "unknown", 
    .default = "n/a" # collective petitions
  ))  |>
  inner_join(bho_hl_urls, by="year") |>
  arrange(year)

```




- [Notes](#notes) on using the Explorer


```{r}
#| column: screen-inset-shaded

###  column: v wide: screen-inset-shaded / not quite so wide: page

reactable(

  data_reactable, 
  
  columns = list(
    petition_id = colDef(show=FALSE), # often won't want to show this sort of id column
    
    # link to BHO page, not to specific entry
    petitioner = colDef(#name="name", 
                        minWidth = 220,
                        sticky = "left",
                        style = sticky_style,
                        headerStyle = sticky_header_style,
                        html = TRUE,
                        # this needs an if... you have to get the url data for it.
                        cell = function(value, index){
                          url_data <- data_reactable$bho_path[index]
                          if (!is.na(url_data)) { sprintf('<a href="https://www.british-history.ac.uk/petitions/%s" target="_blank">%s</a>', data_reactable$bho_path[index], value) } else { value }
                          
                        }
        ), 
    
    year= colDef(name = "year",
                 maxWidth = 120,
                  filterMethod = filterNumPosRange,
                  filterInput = filterTextPlaceholderInput("data-table", "1597/1696"),
                  html=TRUE,
                   # add more info about the column to the header.
                   header = JS('function(column) {
                     return column.name + `<div style="color: #737373; font-weight: normal;">(from/to)</div>`
                   }'),
                  cell = function(value) {
                                if (substr(value, 1,4) == substr(value, 6, 9)) {substr(value, 1, 4)} else {value}
                                }                  
      ),
    
    # handle date as text string.
    date = colDef(name = "date", minWidth = 140 ), 
    
    
    themes = colDef(name="themes"
              #filterInput = selectFilterInput("data-table"),
              # exact text search not needed?
              #filterMethod = filterTextExact
    ),
    
    petition_type = colDef(name="petition type",
               filterInput = selectFilterInput("data-table"),
                # exact text search - i think you might need it here?
               filterMethod = filterTextExact
    ),
    
    # select dropdown + exact text search
    petition_gender = colDef(maxWidth = 100 , 
        filterInput = selectFilterInput("data-table"),
        # exact text search to avoid the 'male' in 'female' problem
        filterMethod = filterTextExact
    ),
    
    
    reference = colDef(name="reference", minWidth = 180),
    
    # columns to hide
    # named_petrs=colDef(show = FALSE),
    # subscribers = colDef(show = FALSE),
    
    bho_path = colDef(show = FALSE),
    description = colDef(show = FALSE)

),

  # groups are esp useful for wider tables, but take a fair bit of work.
  #   columnGroups = list(
  #     colGroup()
  # ),

  elementId = "data-table", # change if more than one on the page!

  filterable = TRUE,

  defaultColDef = colDef(
      header = headerMakeSpace, # replaces underscore with space. often all you need to make headers presentable! 
      minWidth = 120,
      sortNALast = TRUE,
    # start with this commented out as it can slow down large/texty tables. but it's usually worth using.
    ## can exclude specific column(s) by name in the if() eg '& name !="url_text"'
      filterInput = function(values, name) {
        if (is.character(values) ) {
        dataListFilterInput("data-table")(values, name)
      }
    } # /filterInput
   ), # /defaultcoldef


  #defaultColGroup = NULL,

  pagination = TRUE,
  defaultPageSize = 10,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 25, 50, 100),
  paginationType = "numbers", # "numbers" for page number buttons, "jump" for a page jump (more compact up-down arrows, useful for tables with a lot of pages)
  showPageInfo = TRUE,

  
# appearance options might vary a bit 
  highlight = TRUE, # highlight on hover?
  outlined = TRUE, # borders around the table? i quite like the borders actually.
  bordered = TRUE, # borders around the table and every cell?
  borderless = TRUE, #  Remove inner borders from table?
  striped = TRUE, # zebra striping for table rows. how well does this go with borders?
  compact = FALSE, # what diff does this make? I think just a bit less padding etc.

  #groupBy = NULL, # colname or c(colnames). aggregation using aggregate arg in colDef()

  #sortable = TRUE,
  #resizable = FALSE,

  #searchable = FALSE, # won't often want this
  #searchMethod = NULL, # custom search method, JS()
  
  # defaultSortOrder = "asc", # "desc" for descending order
  # defaultSorted = NULL, # colname / c(colnames). for different directions use list() list(Species = "asc", Petal.Length = "desc")

  #minRows = 1,
  #paginateSubRows = FALSE, #  When rows are grouped, paginate sub rows? 
  
  #wrap = TRUE, # text wrapping. If TRUE, long text wrapped to multiple lines. If FALSE, text truncated to fit on one line.
    
  fullWidth = TRUE,

  #meta = NULL, #  Custom metadata to pass to JavaScript render functions or style functions. A named list of values that can also be JS() expressions or functions. 

  defaultExpanded = FALSE, # whether to load with nested rows opened 
  
# details are likely to vary a LOT! 

  # details = NULL, # Additional content to display when expanding a row. An R function that takes the row index and column name as arguments, or a JS() function that takes a row info object as an argument. Can also be a colDef() to customize the details expander column.

#   # You can add details to individual columns! - i think all you need to do is put details inside a colDef rather than at this level.
  details = function(index) {
    det <- data_reactable |>
      select(petition_id, description) |>
      filter(petition_id == data_reactable$petition_id[index])
      # works ok where inquest has no text, but det opens up with "No rows found"; maybe not ideal.
      # maybe https://glin.github.io/reactable/articles/examples.html#conditional-row-details
    tbl <- reactable(det,
                     outlined = TRUE, #highlight = TRUE,
                     #resizable = TRUE, # this does help but it's not immediately obvious that you can resize.
                     #filterable = TRUE, # only filters within the open details field; might be useful if lots of rows.
                     fullWidth = FALSE, # 
                     #defaultPageSize = 30,
                     # possible to use footer for link? not sure how the code would work.
                     columns = list(
                      description=colDef(minWidth=650, html = TRUE ), # can i use % in minWidth? no.
                      petition_id = colDef(show=FALSE)
                    ) # list
                  ) # columns. why isn't a comma needed here???
    htmltools::div(style = list(margin = "10px 60px"), tbl)
  },
  onClick = "expand",
  rowStyle = list(cursor = "pointer")
#   # end of details 

) 
```


## Notes on the Explorer

*Filtering:* 

All columns are filterable.

- Text filters match any part of a cell; just start typing to see a list of possible matches
- The year filter matches ranges using / as separator
  - eg "/1630" = up to 1630, "1630/" = 1630 and later, "1630/1630" = 1630 only, and "1630/1632" = 1630 to 1632
  

*Sorting:*

By default, the table is sorted by year.

- All columns are sortable by clicking on the header
- Sort multiple columns using the Shift key 
- Empty cells are sorted to the *end* of the table



## Notes on the Data

More information about the dataset can be found in the [dataset documentation](https://doi.org/10.5281/zenodo.7027692) and in the British History Online edition's [Introduction](https://www.british-history.ac.uk/series/english-petitions). 

Many petitions are undated and the year has usually been derived from archival context; occasionally, the year might differ slightly from the date.




```{r}
# do some spot checks on these when you have links in the reactable...
# qs_petitions |>
#   left_join(bho_qs_xml_urls, by="petition_id")  |>
#   filter(str_detect(date, "/")  ) |>
#   select(petitioner, reference, date, page) 
```

